LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment

### 2024-07-29 article/main.tex
fix main file
use the review option
fixup some wording

\pdfoutput=1

\documentclass[11pt]{article}
\usepackage[review]{acl}

@@

Firstly, \textbf{is it possible to influence an LLM's writing style by
    instructing it to generate text that complies to certain HLFs}?
This problem of controlled text generation has implications for text style
transfer and authorship verification.
Answering this question might point to cheaper, less intrusive and more user
friendly solutions than fine tuning large language models for these tasks.
Fine tuning LLMs also incurs some loss of generality~\cite{yang2024unveiling},
going directly against the purpose of having a generic language model.

@@

Linguistic style is the sum of identifiable language choices manifest in a text
made from the language system by the text producer~\cite{lugea2023stylistics}.
Another way to think about it is that style is the form used for delivering
meaning~\cite{tst_sigkdd_review_2022}.
Opinions seem to converge that style depends on context and author choice
vis-a-vis a communication goal~\cite{mcdonald1985computational,hovy1987generating}.
The author's choice is available at all linguistic levels, including the
morphological, lexical, syntactic, semantic and pragmatic
levels~\cite{dimarco1994model,lugea2023stylistics}.
The influence of syntactic structure over text style is apparent also from a
formalist perspective~\cite{chomsky2002syntactic}.
Our selection of HLFs is informed by aspects of linguistic style.

@@

In terms of our first question --- whether HLF instructions impact an LLM's
writing, the response is affirmative.
Certainly, not all LLMs are created equal in this regard.
The experiment design involves choices that rely on the presence of a corpus and
the usage of a fixed external input text.
This is relevant as evidentiated by the difference in results obtained using the
Yelp reviews corpus when compared to the results obtained on the CNN article
corpus.
While similarly significant, the HLF results on Yelp reviews are mostly worse
than their corresponding baselines failing to confirm the positive impact of HLF
instructions.
The importance of the choice of the external text is highlighted by the better
results obtained using an input text from the corpus.

This leads us to our second question: how much do HLF instructions impact an
LLM's writing?
The metrics we compute dismiss the idea that HLF instructions have a consistent
desired impact on the generated text.
There isn't conclusive evidence that improving the LLM's odds of producing a
better baseline results in a smaller difference between baseline and result
relative to the target.
More importantly, there is evidence that HLF instructions might have an effect
that is opposite to the desired effect when comparing the results on the Yelp
reviews dataset.

@@

However, the outcomes are not always good.
From a pragmatic standpoint, prompt engineering and a careful choice of language
features and input text seem like a good idea to alleviate results.
Providing examples in the input prompt does not seem as beneficial.

\newpage

@@

\newpage

>>>>

\documentclass[11pt]{article}
\pdfoutput=1

\usepackage[review]{coling}

\usepackage{times}
\usepackage{latexsym}

@@

Firstly, \textbf{is it possible to influence an LLM's writing style by
    instructing it to generate text that complies to certain HLFs}?
This problem of controlled text generation has implications for text style
transfer and authorship verification.
Answering this question might point to cheaper, less intrusive and more user
friendly solutions than fine tuning large language models for these tasks.
Fine tuning LLMs also incurs some loss of generality~\cite{yang2024unveiling},
going directly against the purpose of generic language models.

@@

Linguistic style is the sum of identifiable language choices manifest in a text
made from the language system by the text producer~\cite{lugea2023stylistics}.
Another way to think about it is that style is the form used for delivering
meaning~\cite{tst_sigkdd_review_2022}.
Opinions seem to converge that style depends on context and author choice
vis-a-vis a communication goal~\cite{mcdonald1985computational,hovy1987generating}.
The author's choice is available at all linguistic levels, including the
morphological, lexical, syntactic, semantic and pragmatic
levels~\cite{dimarco1994model,lugea2023stylistics}.
The influence of syntactic structure over text style is apparent also from a
formalist perspective~\cite{chomsky2002syntactic}.
Our selection of HLFs is informed by the various aspects of linguistic style.

@@

In terms of our first question --- whether HLF instructions impact an LLM's
writing, the response is affirmative.
Not all LLMs are equally susceptible to HLF instructions, though.
The experiment design involves choices that rely on the presence of a corpus and
the usage of a fixed external input text.
This is relevant as evidentiated by the difference in results obtained using the
Yelp reviews corpus when compared to the results obtained on the CNN article
corpus.
While similarly significant, the HLF results on Yelp reviews are mostly worse
than their corresponding baselines failing to confirm the positive impact of HLF
instructions.
The importance of the choice of the external text is highlighted by the better
results obtained using an input text from the corpus.

This leads us to our second question: how much do HLF instructions impact an
LLM's writing?
The metrics we compute dismiss the idea that HLF instructions have a consistent
desired impact on the generated text.
There isn't conclusive evidence that improving the LLM's odds of producing a
better baseline results in a smaller difference between baseline and result
relative to the target.
More importantly, there is evidence that HLF instructions have an adverse effect
when comparing the results on the Yelp reviews dataset.

@@

However, the outcomes are not always good.
From a pragmatic standpoint, prompt engineering and a careful choice of language
features and input text seem like the way to obtain desirable results.
Providing examples in the input prompt does not seem as beneficial.

@@
